\documentclass[a4paper,oneside,11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[light,nott]{kpfonts}
\usepackage[left=30mm,right=30mm,top=30mm,bottom=40mm]{geometry}
\usepackage{tabularx}

\usepackage{lipsum}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{framed}
\usepackage[%
        bookmarksnumbered=true,
        colorlinks=true,
        linkcolor=cyan!50!blue,
        citecolor=violet,
        urlcolor=purple,
    ]{hyperref}
\usepackage{hologo}

\newcommand\compar[2]{\texttt{\textbackslash #1\{#2\}}}
\newcommand\comm[1]{\texttt{\textbackslash#1}}
\newcommand\code[1]{\texttt{#1}}
\newcommand\XeLaTeX{\hologo{XeLaTeX}}
\newenvironment{framedcompact}
    {\begin{framed}\vspace{-1em}}
    {\vspace{-1em}\end{framed}}
\definecolor{TFFrameColor}{HTML}{4393c3}
\definecolor{TFTitleColor}{HTML}{FFFFFF}

\title{Changing Fonts in \LaTeX}
\author{Raphael Frey  <\href{mailto:webmaster@alpenwasser.net}{\nolinkurl{webmaster@alpenwasser.net}}>}


% ************************************************************************** %
\begin{document}
% ************************************************************************** %

\maketitle

\vspace{4em}

\begin{abstract}
    This document  presents a  very brief  overview of the  topic of  fonts in
    \LaTeX. A  quick introduction  to some  of the  terminology is  given, and
    various ways  are shown to change  fonts and different aspects  of them to
    one's liking.

    This document is in  no way, shape or form exhaustive;  the topic of fonts
    can fill  (and has filled)  entire books. But  it should suffice  for most
    day-to-day usage,  and where it does  not, provides some pointers  to more
    in-depth resources.
\end{abstract}

\vspace{4em}

\tableofcontents


% -------------------------------------------------------------------------- %
\newpage
\section{Incomplete Summary For the Impatient}
\label{sec:summary}
% -------------------------------------------------------------------------- %

\begin{titled-frame}
{\textsf{Switching to Default Document Font}}
\vspace{-1em}
\begin{verbatim}
\normalfont The quick brown fox jumps over the lazy dog.
\end{verbatim}
\normalfont The quick brown fox jumps over the lazy dog.
\end{titled-frame}

\vspace{-1.0em}

\begin{titled-frame}
{\textsf{Changing Fonts between Roman, Sans Serif and Monospace}}
\vspace{-1em}
\begin{verbatim}
\rmfamily We shall start with roman, \sffamily continue with
sans serif, \ttfamily and end up at typewriter.
\end{verbatim}
\rmfamily  We shall  start with  roman,  \sffamily continue  with sans  serif,
\ttfamily and end up at typewriter.
\end{titled-frame}

\vspace{-1.0em}

\begin{titled-frame}
{\textsf{Writing a Piece of Text In Roman, Sans Serif and Monospace}}
\vspace{-1em}
\begin{verbatim}
\textrm{The quick brown fox jumps over the lazy dog.}
\textsf{The quick brown fox jumps over the lazy dog.}
\texttt{The quick brown fox jumps over the lazy dog.}
\end{verbatim}
\textrm{The quick brown fox jumps over the lazy dog.}\\
\textsf{The quick brown fox jumps over the lazy dog.}\\
\texttt{The quick brown fox jumps over the lazy dog.}
\end{titled-frame}

\vspace{-1.0em}

\begin{titled-frame}
{\textsf{Writing a Piece of Text in Different Font Weights}}
\vspace{-1em}
\begin{verbatim}
\textbf{The quick brown fox jumps over the lazy dog.}
\textmd{The quick brown fox jumps over the lazy dog.}
\end{verbatim}
\textbf{The quick brown fox jumps over the lazy dog.}\\
\textmd{The quick brown fox jumps over the lazy dog.}
\end{titled-frame}

\vspace{-1.0em}

\begin{titled-frame}
{\textsf{Writing a Piece of Text in Different Font Shapes}}
\vspace{-1em}
\begin{verbatim}
\emph{The quick brown fox jumps over the lazy dog.}
\textit{The quick brown fox jumps over the lazy dog.}
\textsl{The quick brown fox jumps over the lazy dog.}
\textsc{The quick brown fox jumps over the lazy dog.}
\textup{The quick brown fox jumps over the lazy dog.}
\end{verbatim}
\emph{The quick brown fox jumps over the lazy dog.}\\
\textit{The quick brown fox jumps over the lazy dog.}\\
\textsl{The quick brown fox jumps over the lazy dog.}\\
\textsc{The quick brown fox jumps over the lazy dog.}\\
\textup{The quick brown fox jumps over the lazy dog.}
\end{titled-frame}


% -------------------------------------------------------------------------- %
\newpage
\section{Terminology: Font Families, Fonts and Glyphs}
\label{sec:font-families}
% -------------------------------------------------------------------------- %

A \emph{font family},  also referred to as  a \emph{typeface}, is a  font or a
set  of fonts  with common  features and  characteristics, comprising  all the
needed \emph{glyphs}. A glyph, in this  context, is a graphical representation
of a character,  which can be a  letter, a logogram, a  mathematical symbol or
other things\footnotemark.

\footnotetext{%
    Obviously typography is a vast and complex topic and there is no way I can
    do it  true justice  here. Even if  I wanted  to, I  am most  certainly no
    scholar  on  the  topic. But  for  the  purposes  of  this  document  this
    information shall suffice. The curious reader  shall feel free to research
    the topic in more depth, if so desired.}

For  example, Table  \ref{tab:allographs}  contains  different glyphs  (called
allographs) of the letter Q.

\begin{table}[h!]
    \centering
    \caption{Allographs for the letter Q}
    \label{tab:allographs}
    \vspace{0.5em}
    \begin{tabular}{llr}
        \toprule
        Font                               & \TeX{} Abbreviation & Glyph \\
        \midrule
        Computer Modern (\LaTeX{} default) & \code{cmr}  & \fontfamily{cmr}\selectfont Q \\
        Latin Modern                       & \code{lmr}  & \fontfamily{lmr}\selectfont Q \\
        Johannes Kepler Roman              & \code{jkp}  & \fontfamily{jkp}\selectfont Q\\
        Johannes Kepler Roman Light        & \code{jkpl} & \fontfamily{jkpl}\selectfont Q\\
        Latin Modern Sans Serif            & \code{lmss} & \fontfamily{lmss}\selectfont Q \\
        Avant Garde                        & \code{pag}  & \fontfamily{pag}\selectfont Q \\
        Computer Modern Typewriter         & \code{cmtt} & \fontfamily{cmtt}\selectfont Q \\
        Zapf Chancery                      & \code{pzc}  & \fontfamily{pzc}\selectfont Q \\
        \bottomrule
    \end{tabular}
\end{table}

In general,  font families  which one  tends to use  in a  reasonable document
these days tend to be grouped into three main categories\footnotemark:

\footnotetext{%
    There are  more, but  these are  unlikely to be  useful in  most practical
    applications these days. For more, see \cite{wikipedia:typeface}.}

\begin{titled-frame}
    {\textsf{Common Font Families}}
    \noindent{\fontfamily{ppl}\selectfont roman fonts (with serifs), for example Palatino}\\
    {\fontfamily{phv}\selectfont sans-serif fonts, for example Helvetica}\\
    {\fontfamily{pcr}\selectfont monospace fonts, for example Courier}
\end{titled-frame}

Often, font families  provide fonts for roman, sans-serif  and monospace fonts
(also called  typewriter fonts,  particularly in  the \TeX{}  world). In those
cases, the  metrics and aesthetics of  these different fonts in  the same font
family have  been tuned to look  pleasing when used together. It  is therefore
often tricky to  mix and match fonts from different  font families without the
result looking odd\footnotemark.

\footnotetext{%
    Of course it  can be done, but it  is helpful to be aware  of this general
    issue and to  be careful when picking fonts from  different families to be
    used together in the same document.}

In \LaTeX, there are three macros which contain the default fonts families for
the three font groups:

\begin{titled-frame}
    {\textsf{Default Font Families in \LaTeX}}
    \noindent\comm{rmdefault}\\
    \comm{sfdefault}\\
    \comm{ttdefault}
\end{titled-frame}

Furthermore,  there  is  a  command \comm{familydefault}  which  contains  the
currently configured default font family (one of the above three).


% -------------------------------------------------------------------------- %
\section{Selecting Fonts}
\label{sec:setting-fonts}
% -------------------------------------------------------------------------- %

Fundamentally,  there  are  three  mechanisms  for  selecting  fonts: Firstly,
\emph{font     declarations}     and,    secondly,     their     corresponding
\emph{commands}. Declarations change the font  within the current scope, while
the  commands only  typeset the  content within  their argument,  delimited by
\verb|{| and \verb|}|. Table \ref{tab:font-decs-cmds} presents an overview the
declarations, their equivalent commands and their effect.

\begin{table}[h!]
    \centering
    \caption{Font declarations and commands, and their effects}
    \vspace{0.5em}
    \label{tab:font-decs-cmds}
    \begin{tabular}{>{\ttfamily}l<{\normalfont} >{\ttfamily}l<{\normalfont} l}
        \toprule
        \normalfont\scshape Declaration & \normalfont\scshape Command  & \scshape Effect \\
        \midrule
        \textbackslash{}mdseries   & \textbackslash{}textmd\{\textit{text}\}     & \mdseries Medium Series             \\
        \textbackslash{}bfseries   & \textbackslash{}textbf\{\textit{text}\}     & \bfseries Bold Series               \\
        \textbackslash{}upshape    & \textbackslash{}textup\{\textit{text}\}     & \upshape Upright Shape              \\
        \textbackslash{}itshape    & \textbackslash{}textit\{\textit{text}\}     & \itshape Italic Shape               \\
        \textbackslash{}slshape    & \textbackslash{}textsl\{\textit{text}\}     & \slshape Slanted Shape              \\
        \textbackslash{}scshape    & \textbackslash{}textsc\{\textit{text}\}     & \scshape Small Caps Shape           \\
        \textbackslash{}scslshape  & \textbackslash{}textscsl\{\textit{text}\}   & \scslshape Small Caps Slanted Shape \\
        \textbackslash{}rmfamily   & \textbackslash{}rmfamily\{\textit{text}\}   & \rmfamily Roman Family              \\
        \textbackslash{}sffamily   & \textbackslash{}sffamily\{\textit{text}\}   & \sffamily Sans Serif Family         \\
        \textbackslash{}ttfamily   & \textbackslash{}ttfamily\{\textit{text}\}   & \ttfamily Typewriter Family         \\
        \textbackslash{}normalfont & \textbackslash{}normalfont\{\textit{text}\} & \normalfont Default Document Font   \\
        \bottomrule
    \end{tabular}
\end{table}

Different fonts  may or may not  implement some or more  of these options. For
example, the  \textscsl{Small Caps Slanted}  font variant is not  supported by
Computer Modern, but comes with the Kp-Fonts font family which has been chosen
for this document.

The  third mechanism  is to  set the  four properties  which are  configurable
for  a  font  with   the  commands  \verb|\fontencoding|,  \verb|\fontfamily|,
\verb|\fontseries| and \verb|\fontshape|,  followed by \verb|\selectfont| (not
all of  them need be set,  of course; one may  choose to only change  the font
family, for example). The  \verb|\selectfont| command tends to  be used mostly
for local font style  changes which can't be covered by  the macros from Table
\ref{tab:font-decs-cmds}.


\begin{titled-frame}
{\textsf{The \textbackslash{}selectfont Command}}
\vspace{-1em}
\begin{verbatim}
\fontencoding{T1}
\fontfamily{pag}
\fontseries{m}
\fontshape{it}
\selectfont
The quick brown fox jumps over the lazy dog.
\end{verbatim}
\fontencoding{T1}
\fontfamily{pag}
\fontseries{m}
\fontshape{it}
\selectfont
The quick brown fox jumps over the lazy dog.
\end{titled-frame}


% -------------------------------------------------------------------------- %
\subsection{Font Packages}
\label{subsec:font-packages}
% -------------------------------------------------------------------------- %

The most convenient way to choose the default fonts for a document is by using
font packages. Once  one or more font  packages have been loaded,  we can then
switch among the  defaults which have been defined by  those packages with the
commands from Table \ref{tab:font-decs-cmds}.

Good places  to peruse for  finding fonts  are CTAN \cite{ctan:fonts}  and the
\LaTeX{}  font  catalogue  \cite{tug:font-catalog}. A popular  choice  is  the
Latin Modern  Package (\verb|\usepackage{lmodern}|), which is  an expansion of
Computer Modern  with better  support for non-ASCII  characters. For documents
which are not in  English, Latin Modern is a solid  choice, provided one likes
its aesthetics.

As  a  different example,  if  we  wish to  typeset  our  document in  DejaVu,
we  can  put  \compar{usepackage}{dejavu}{}  in the  preamble. This  will  set
\comm{rmdefault}, \comm{sfdefault}  and \comm{ttdefault} to  the corresponding
values, while \comm{familydefault} will continue to  point to the one of those
three to which it was set previously (\comm{rmdefault} by default).

Not all font packages provide all kinds of font families, or all font families
in all  weights and  shapes. Consult the documentation  for the  package which
you're intending to use. In the cases  where a package only provides some type
of font, the  others will be left  unchanged and may be configured  by using a
font  package  specific  to  those  fonts, or  left  at  the  Computer  Modern
default. For example, the  \code{FiraSans} package, which is  obviously a sans
serif typeface, does not provide a  roman font. Therefore, the serif fonts are
left untouched by loading that package.

The same  goes for mathematics: Not all  fonts provide the needed  symbols for
typesetting mathematics.

A personal  favorite of mine  is the Kp-Fonts family  \cite{ctan:kpfonts}.  It
has  roman, sans-serif  and  typewriter font  choices  (although personally  I
prefer  the Computer  Modern typewriter  font and  usually keep  that intact),
along with  mathematics and  many symbols. This  is the  setup which  has been
chosen for this  document. But this is a matter of  personal preference; one's
mileage may vary.


% -------------------------------------------------------------------------- %
\subsection{Global Font Selection For a Document}
\label{subsec:fontsel:global}
% -------------------------------------------------------------------------- %

One may wish to typeset a  document in sans serif or typewriter. Besides using
a package  to do this,  one may also  manually configure the  default document
font in the preamble\footnotemark like this:

\footnotetext{%
    Or anywhere else in the document,  for that matter. Whether or not that is
    a sensible  thing to do  is another  question. One should usually  not mix
    different fonts and font styles willy-nilly.}

\begin{titled-frame}
{\textsf{Defining the Default Font Families Manually}}
\vspace{-1em}
\begin{verbatim}
\renewcommand{\familydefault}{\rmdefault}
\renewcommand{\familydefault}{\sfdefault}
\renewcommand{\familydefault}{\ttdefault}
\end{verbatim}
\vspace{-1em}
\end{titled-frame}

The first  command will  usually do  nothing, of  course, because  the default
family tends to be the roman font.

Note that typewriter fonts differ in  what kinds of adjustments \TeX{} usually
makes  when it  comes to  spacing in  order to  achieve a  justified block  of
text.  Also,  words in typerwriter fonts  will not hyphenate except  in places
where  \code{\textbackslash{}-}  has  been  inserted  manually. This  is  also
why  sometimes a  monospaced  piece of  text can  protrude  outside the  right
textblock margin  when you put  it in normal text  (see above with  the dejavu
example; that  was not  done on purpose). More  on the topic  can be  found in
\cite{texblog:typewriter}.

Combining different font families can be  done in the same manner. This allows
to compose  a custom set of  default fonts for  a document which might  not be
covered by  a font package. As mentioned,  care should be taken  to pick fonts
which go nicely together. For example, selecting Palatino as the roman default
font,  Helvetica as  the  default  sans-serif font  and  Latin  Modern as  the
typewriter font could be done by putting these commands in the preamble:

\begin{titled-frame}
{\textsf{Creating a Custom Set of Default Fonts}}
\vspace{-1em}
\begin{verbatim}
\renewcommand{\rmdefault}{ppl}  % Palatino
\renewcommand{\sfdefault}{phv}  % Helvetica
\renewcommand{\ttdefault}{lmtt} % Latin Modern
\end{verbatim}
\vspace{-1em}
\end{titled-frame}

Changing  math  fonts is  rather  more  complex. I  recommend not  doing  that
manually,  but  instead  relying  on   font  packages  instead,  as  described
in  the  previous  section. See for  example  \cite{stackexch:math-fonts}  and
\cite{practex:fonts}).


% -------------------------------------------------------------------------- %
\subsection{Local Font Selection From Defined Defaults}
\label{subsec:fontsel:local:defaults}
% -------------------------------------------------------------------------- %

There  are  two  basic  scenarios  for  locally  changing  fonts: Selecting  a
different font, series  or shape from the defined default  font families which
are already loaded, or selecting a completely different font which has not yet
been loaded  into the document. This  section deals with the  former scenario;
for the latter, see the next section.

Locally  changing  font  properties  or families  among  the  choices  already
loaded  can  be done  with  the  declarations  and  commands listed  in  Table
\ref{tab:font-decs-cmds}.


\begin{titled-frame}
{\textsf{Local Font Changing Among Loaded Fonts and Styles}}
\vspace{-1em}
\begin{verbatim}
\bfseries This text is in bold, while \mdseries this text is
in  normal weight.  \slshape  We continue  with the  slanted
shape,  followed  by  the  \scshape  small  capitals  shape,
followed by  the \itshape italic shape.  \normalfont and now
we are back to the  default document font, whatever that may
be.

{\scslshape For keeping font  declaration effects local}, we
must enclose  the declaration and  the text which  it should
affect inside a scope.
\end{verbatim}
\bfseries This text is in bold, while \mdseries this text is
in  normal weight.  \slshape  We continue  with the  slanted
shape,  followed  by  the  \scshape  small  capitals  shape,
followed by  the \itshape italic shape.  \normalfont And now
we are back to the  default document font, whatever that may
be.

{\scslshape For keeping font  declaration effects local}, we
must enclose  the declaration and  the text which  it should
affect inside a scope.
\end{titled-frame}

Or with font commands:

\begin{titled-frame}
{\textsf{Local Font Changing Among Loaded Fonts and Styles}}
\vspace{-1em}
\begin{verbatim}
\textbf{This text is in bold, while} \textmd{this text is in
normal weight}. \textsl{We continue  with the slanted shape,
followed by the } \textsc{ small capitals shape, followed by
the  }\textit{italic  shape.} And now  we  are  back to  the
default document font, whatever that may be.
\end{verbatim}
\textbf{This text is in bold, while} \textmd{this text is in
normal weight}. \textsl{We continue  with the slanted shape,
followed by the } \textsc{ small capitals shape, followed by
the  }\textit{italic  shape.} And now  we  are  back to  the
default document font, whatever that may be.
\end{titled-frame}


% -------------------------------------------------------------------------- %
\subsection{Local Font Selection From Non-Defaults}
\label{subsec:fontsel:local:nondefaults}
% -------------------------------------------------------------------------- %

If we  wish to  load a completely  different font which  has not  already been
loaded (be  that because we  did not  want it to  be the document  default, or
because  there simply  does  not exist  a  package  for it),  we  can use  the
\verb|\selectfont| command and its friends.  

\begin{titled-frame}{\sffamily Switching to a Non-Default Font}
For example, if we wish to switch to Helvetica:

\vspace{1em}\noindent
\verb|\fontfamily{phv}\selectfont|\\

\fontfamily{phv}\selectfont  And  now  we   should  have  Helvetica  for  this
text. The font family  will stay Helvetica until it is  switched back with one
of  these two  commands  (or changed  to anything  else  with the  appropriate
command, should one so choose):

\vspace{1em}\noindent
\verb|\fontfamily{\familydefault}\selectfont|\\[1ex]
or\\[1ex]
\noindent
\verb|\normalfont|\\

\noindent\normalfont And now we're back to Kp-Fonts serif.
\end{titled-frame}

The same scope rules which we've already seen apply here as well:

\begin{titled-frame}
{\textsf{Selecting Arbitrary Font Families Locally}}
\vspace{-1em}
\begin{verbatim}
{\fontfamily{phv}\selectfont This text is in Helvetica!}
While this text is not!
\end{verbatim}
{\fontfamily{phv}\selectfont This text is in Helvetica!} While this text is not!
\end{titled-frame}


% -------------------------------------------------------------------------- %
\section{Font Properties}
\label{sec:font-props}
% -------------------------------------------------------------------------- %

We  will  now look  at  changing  various  properties  for a  font: Its  size,
its  series (weight)  and  its  shape. For a  nice  explanation  of this,  see
\cite{stackexch:bfseries-textbf}.


% -------------------------------------------------------------------------- %
\subsection{Font Sizes}
\label{subsec:font-sizes}
% -------------------------------------------------------------------------- %

By  default, \LaTeX{}  offers the  following font  sizes, which,  depending on
the  global  document  font  size  set in  the  preamble  (for  example,  with
\code{\textbackslash{}documentclass[11pt]\{article\}}), vary in the size which
they will have on the page:

\begin{center}
    \begin{tabular}{lrrrrrr}
        \toprule
        \textsc{Size} &
        \multicolumn{3}{l}{\textsc{Standard Classes and beamer}} &
        \multicolumn{3}{l}{\textsc{AMS Classes and memoir}} \\
        \midrule
        \comm{tiny}         &  5    &  6     & 6     & 6     & 7     & 8     \\
        \comm{scriptsize}   &  7    &  8     & 8     & 7     & 8     & 9     \\
        \comm{footnotesize} &  8    &  9     & 10    & 8     & 9     & 10    \\
        \comm{small}        &  9    &  10    & 10.95 & 9     & 10    & 10.95 \\
        \comm{normalsize}   & 10    &  10.95 & 12    & 10    & 10.95 & 12    \\
        \comm{large}        & 12    &  12    & 14.4  & 10.95 & 12    & 14.4  \\
        \comm{Large}        & 14.4  &  14.4  & 17.28 & 12    & 14.4  & 17.28 \\
        \comm{LARGE}        & 17.28 &  17.28 & 20.74 & 14.4  & 17.28 & 20.74 \\
        \comm{huge}         & 20.74 &  20.74 & 24.88 & 17.28 & 20.74 & 24.88 \\
        \comm{Huge}         & 24.88 &  24.88 & 24.88 & 20.74 & 24.88 & 24.88 \\
        \bottomrule
    \end{tabular}
\end{center}

The table has been taken from \cite{wikibooks:fonts}.

Yet again, these commands change the font size within the current scope, so should
usually be enclosed in braces:

\begin{titled-frame}
{\textsf{Locally Changing Font Sizes}}
\vspace{-1em}
\begin{verbatim}
{\LARGE This is a piece of large text.} And this is of normal size again.
\end{verbatim}
{\LARGE This is a piece of large text.} And this is of normal size again.
\end{titled-frame}

% -------------------------------------------------------------------------- %
\subsection{Font Series}
\label{subsec:font-series}
% -------------------------------------------------------------------------- %

Changing font weights can be done either via a \comm{fontseries\{<argument>\}}
command,   followed  by   \comm{selectfont},  or   via  the   \comm{bfseries},
\comm{mdseries}   and    \comm{lfseries}   declarations   listed    in   Table
\ref{tab:font-decs-cmds},  respectively. A  font  may offer  more  than  three
weights,  in which  case one  might need  to resort  to the  \comm{fontseries}
command  for   changing  fonts   because  the   corresponding  \comm{..series}
declaration does not exist.

Furthermore,  depending on  which option  is  loaded if  one is  using a  font
package  such as  Kp-Fonts, the  default font  weight and  all others  will be
changed accordingly  among \comm{bfseries} and its  friends. For example, this
happens when  one selects  the \comm{light}  option for  Kp-Fonts package. For
specifics, consult the documentation for the package you're using.

Lastly,  note that  not  all options  might  be supported  by  all fonts  (for
example,  Kp-Fonts  with  a  \code{light} option  shows  vehement  displeasure
at  being  instructed  to   execute  \comm{lfseries}). With  Kp-Fonts  in  the
configuration for this document, these are the options available:

\begin{titled-frame}
{\textsf{Locally Changing Font Weight}}
\vspace{-1em}
\begin{verbatim}
{\fontseries{l}\selectfont  The quick brown fox jumps over the lazy dog.}
{\fontseries{m}\selectfont  The quick brown fox jumps over the lazy dog.}
{\fontseries{sb}\selectfont The quick brown fox jumps over the lazy dog.}
{\fontseries{b}\selectfont  The quick brown fox jumps over the lazy dog.}
{\fontseries{bx}\selectfont The quick brown fox jumps over the lazy dog.}

{\mdseries\selectfont The quick brown fox jumps over the lazy dog.}
{\bfseries\selectfont The quick brown fox jumps over the lazy dog.}

\textbf{The quick brown fox jumps over the lazy dog.}
\textmd{The quick brown fox jumps over the lazy dog.}
\end{verbatim}

\noindent{\fontseries{l}\selectfont  The quick brown fox jumps over the lazy dog.}\\
{\fontseries{m}\selectfont  The quick brown fox jumps over the lazy dog.}\\
{\fontseries{sb}\selectfont  The quick brown fox jumps over the lazy dog.}\\
{\fontseries{b}\selectfont  The quick brown fox jumps over the lazy dog.}\\
{\fontseries{bx}\selectfont The quick brown fox jumps over the lazy dog.}\\

\noindent{\mdseries\selectfont The quick brown fox jumps over the lazy dog.}\\
{\bfseries\selectfont The quick brown fox jumps over the lazy dog.}\\

\noindent\textmd{The quick brown fox jumps over the lazy dog.}\\
\textbf{The quick brown fox jumps over the lazy dog.}\\
\vspace{-1em}
\end{titled-frame}


% -------------------------------------------------------------------------- %
\subsection{Font Shapes}
\label{subsec:font-shapes}
% -------------------------------------------------------------------------- %

Selecting the  font shape can  be done via  the \comm{fontshape\{<argument>\}}
command,   followed   by   \comm{selectfont},    by   a   corresponding   font
declaration,  or, by  the  corresponding \verb|\text..{}|  command (see  Table
\ref{tab:font-decs-cmds}).   Again, the  number of  font shapes   and possible
combinations  thereof will  vary depending  on the  font being  used. Kp-Fonts
supports these options:

\begin{titled-frame}
{\textsf{Locally Changing Font Shapes}}
\vspace{-1em}
\small
\begin{verbatim}
{\fontshape{sc}\selectfont The quick brown fox jumps over the lazy dog.}
{\fontshape{scsl}\selectfont The quick brown fox jumps over the lazy dog.}
{\fontshape{it}\selectfont The quick brown fox jumps over the lazy dog.}
{\otherscshape\selectfont The quick brown fox jumps over the lazy dog.}
{\otherscslshape\selectfont The quick brown fox jumps over the lazy dog.}

{\upshape The quick brown fox jumps over the lazy dog.}
{\itshape The quick brown fox jumps over the lazy dog.}
{\slshape The quick brown fox jumps over the lazy dog.}
{\scshape The quick brown fox jumps over the lazy dog.}
{\scslshape The quick brown fox jumps over the lazy dog.}

\textup{The quick brown fox jumps over the lazy dog.}
\textit{The quick brown fox jumps over the lazy dog.}
\textsl{The quick brown fox jumps over the lazy dog.}
\textsc{The quick brown fox jumps over the lazy dog.}
\textscsl{The quick brown fox jumps over the lazy dog.}

\end{verbatim}
\normalsize
\noindent{\fontshape{sc}\selectfont The quick brown fox jumps over the lazy dog.}\\
{\fontshape{scsl}\selectfont The quick brown fox jumps over the lazy dog.}\\
{\fontshape{it}\selectfont The quick brown fox jumps over the lazy dog.}\\
{\otherscshape\selectfont The quick brown fox jumps over the lazy dog.}\\
{\otherscslshape\selectfont The quick brown fox jumps over the lazy dog.}\\\\
{\upshape The quick brown fox jumps over the lazy dog.}\\
{\itshape The quick brown fox jumps over the lazy dog.}\\
{\slshape The quick brown fox jumps over the lazy dog.}\\
{\scshape The quick brown fox jumps over the lazy dog.}\\
{\scslshape The quick brown fox jumps over the lazy dog.}\\\\
\textup{The quick brown fox jumps over the lazy dog.}\\
\textit{The quick brown fox jumps over the lazy dog.}\\
\textsl{The quick brown fox jumps over the lazy dog.}\\
\textsc{The quick brown fox jumps over the lazy dog.}\\
\textscsl{The quick brown fox jumps over the lazy dog.}\\
\vspace{-1em}
\end{titled-frame}


% -------------------------------------------------------------------------- %
\subsection{Combinations}
\label{subsec:combinations}
% -------------------------------------------------------------------------- %

The  commands presented  above can  be  combined in  various ways,  depending,
again,  on the  font families  being used. Some  fonts will  support different
combinations than others.

\begin{titled-frame}
{\textsf{Combining Different Font Styles}}
\vspace{-1em}
\begin{verbatim}
The quick brown fox jumps over the lazy dog.
{\fontshape{sc}\fontseries{b}\selectfont
The quick brown fox jumps over the lazy dog.}
{\fontshape{it}\fontseries{b}\selectfont
The quick brown fox jumps over the lazy dog.}
\end{verbatim}

\noindent The quick brown fox jumps over the lazy dog.\\
{\fontshape{sc}\fontseries{b}\selectfont
The quick brown fox jumps over the lazy dog.}\\
{\fontshape{it}\fontseries{b}\selectfont
The quick brown fox jumps over the lazy dog.}
\end{titled-frame}

% -------------------------------------------------------------------------- %
\subsection{Emphasis}
% -------------------------------------------------------------------------- %

The  \comm{emph}  command  is  a   context-aware  way  to  highlight  text. It
will  select  an  appropriate  font  version  automatically  to  achieve  this
goal. Usually, \comm{emph} uses  italics, but when it is  placed inside italic
text, it will switch to an upright font shape.

\begin{titled-frame}
{\textsf{Emphasis}}
\vspace{-1em}
\small
\begin{verbatim}
{\fontshape{u}\selectfont This sentence \emph{emphasizes} a word of itself.}
{\fontshape{it}\selectfont This sentence \emph{emphasizes} a word of itself.}
Here, again, there exists an {\em alternative} version.
{\itshape Here, again, there exists an {\em alternative} version.}
\end{verbatim}
\normalsize
{\fontshape{u}\selectfont This sentence \emph{emphasizes} a word of itself.}\\
{\fontshape{it}\selectfont This sentence \emph{emphasizes} a word of itself.}\\
Here, again, there exists an {\em alternative} version.\\
{\itshape Here, again, there exists an {\em alternative} version.}
\end{titled-frame}

%% -------------------------------------------------------------------------- %
%\section{List of Typefaces}
%\label{sec:tflist}
%% -------------------------------------------------------------------------- %
%
%\begin{itemize}
%    \item
%        {\fontfamily{cmr}\selectfont Computer Modern: The \LaTeX{} default}
%    \item
%        {\fontfamily{lmr}\selectfont Latin Modern: A  modernised derivative of
%        Computer modern with more characters.}
%    \item
%        {\fontfamily{jkp}\selectfont Kp-Fonts: An alternative to the above two
%        with fully-featured  glyph sets for roman,  sans-serif, typewriter and
%        math fonts. \cite{ctan:kpfonts}}
%    \item
%        {\fontfamily{aeccr}\selectfont aecc --  Almost European Concrete Roman
%        virtual  fonts.  A  font prividing  roman, sans-serif,  and typewriter
%        fonts. \cite{ctan:aecc}}
%\end{itemize}



% -------------------------------------------------------------------------- %
\newpage
\begin{thebibliography}{1}
% -------------------------------------------------------------------------- %

    \bibitem{wikipedia:typeface}
        Wikipedia,
        ``\emph{Typeface}'',
        [Online],
        \href{https://en.wikipedia.org/wiki/Typeface}
             {\nolinkurl{https://en.wikipedia.org/wiki/Typeface}},
        [Accessed: 2017-MAR-20].

    \bibitem{texblog:typewriter}
        Stefan,
        ``\emph{Full justification with typewriter font}'',
        [Online],
        \href{http://texblog.net/latex-archive/plaintex/full-justification-with-typewriter-font/}
             {\nolinkurl{http://texblog.net/latex-archive/plaintex/full-justification-with-typewriter-font/}},
        2008-MAY-08,
        [Accessed: 2017-MAR-20].

    \bibitem{ctan:fonts}
        Comprehensive \TeX{} Archive Network,
        ``\emph{Topic Font: Fonts Themselves}'',
        [Online],
        \href{http://ctan.org/topic/font}
             {\nolinkurl{http://ctan.org/topic/font}},
        [Accessed: 2017-MAR-20].

    \bibitem{ctan:kpfonts}
        The Johannes Kepler Project, Christophe Caignaert,
        ``\emph{kpfonts -- A complete set of fonts for text and mathematics}''
        [Online],
        \href{https://www.ctan.org/pkg/kpfonts}
             {nolinkurl{https://www.ctan.org/pkg/kpfonts}},
        [Accessed: 2017-MAR-22].

    \bibitem{tug:font-catalog}
        \TeX{} Users Group,
        ``\emph{The \LaTeX{} Font Catalogue}'',
        [Online],
        \href{http://www.tug.dk/FontCatalogue/}
             {\nolinkurl{http://www.tug.dk/FontCatalogue/}},
        [Accessed: 2017-MAR-20].

    \bibitem{stackexch:math-fonts}
        tex.stackexchange.com,
        ``\emph{how to select math font in document}'',
        [Online],
        \href{http://tex.stackexchange.com/questions/30049/how-to-select-math-font-in-document}
             {\nolinkurl{http://tex.stackexchange.com/questions/30049/how-to-select-math-font-in-document}},
        [Accessed: 2017-MAR-20].

    \bibitem{practex:fonts}
        Walter Schmidt,
        ``\emph{Font selection in \LaTeX{}: The most frequently asked questions}'',
        The Prac\TeX{} Journal,
        [Online],
        \href{https://www.tug.org/pracjourn/2006-1/schmidt/schmidt.pdf}
             {\nolinkurl{https://www.tug.org/pracjourn/2006-1/schmidt/schmidt.pdf}},
        2006-FEB-01,
        [Accessed: 2017-MAR-20].

    \bibitem{wright:font-commands}
        David Wright,
        ``\emph{\LaTeX{} Font Commands}'',
        [Online],
        \href{http://www.cl.cam.ac.uk/~rf10/pstex/latexcommands.htm}
             {\nolinkurl{http://www.cl.cam.ac.uk/~rf10/pstex/latexcommands.htm}},
        1998,
        [Accessed: 2017-MAR-20].

    \bibitem{stackexch:bfseries-textbf}
        Tobi,
        ``\emph{\comm{bfseries} is to \comm{textbf} as WHAT is to \comm{textsf}}'',
        [Online],
        \href{http://tex.stackexchange.com/a/139592}
             {\nolinkurl{http://tex.stackexchange.com/a/139592}},
        2013-OCT-19,
        [Accessed: 2017-MAR-21].

    \bibitem{wikibooks:fonts}
        wikibooks,
        ``\emph{\LaTeX/Fonts}'',
        [Online],
        \href{https://en.wikibooks.org/wiki/LaTeX/Fonts}
             {\nolinkurl{https://en.wikibooks.org/wiki/LaTeX/Fonts}},
        [Accessed: 2017-MAR-21].

%    \bibitem{ctan:aecc}
%        CTAN,
%        ``\emph{aecc -- Almost European Concrete Roman virtual fonts}'',
%        [Online],
%        \href{http://ctan.org/pkg/aecc}
%             {\nolinkurl{http://ctan.org/pkg/aecc}},
%        [Accessed: 2017-MAr-22].

\end{thebibliography}

\end{document}
